Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-08-02 ms_pre_post.Rmd
Tweaks to methods to make clearer

In the context of inbreeding as the postulated driver of reproductive strategy evolution, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will be affected not only by the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also by the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Empirical studies on diverse systems have demonstrated diverse costs of mating and mate choice, for example, including energetic costs of developing, maintaining, or enacting necessary physiologies [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]; increased risks of predation or disease stemming from increased mate-searching or mating [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]; increased risk of complete mating or fertilisation failure given extreme choosiness [@Kokko2013]; and risks of harm stemming from sexual conflict over fertilisation [e.g., @Rowe1994]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might go extinct, especially if their effects become redundant following evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the inbreeding avoidance benefit of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple potentially costly inbreeding avoidance strategies that might influence each other's evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

@@

Phenotypic values of the three reproductive strategy traits each incur set costs that combine independently to increase the probability that a focal female will die before mating. Numerous forms and mechanisms of direct costs of reproductive strategies could potentially be hypothesised and modelled; the most appropriate formulation depends on the question posed (see Discussion). By allowing costs of the three strategies to be directly and independently controlled, our model facilitates direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance given known cost asymmetries. Such costs on female survival probably are biologically reasonable. For example, polyandrous females that undertake increased mate searching or mating can experience increased predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice can increase the risk of mortality due to harm caused by sexual conflict over mating, and also risk complete mating failure [which equates to pre-reproductive mortality in semelparous organisms; @Rowe1994; @Kokko2013]. Finally, females that express post-copulatory choice can pay up-front energetic costs, and associated trade-offs with survival, due to developing physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014].

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs (see Supporting Information).

>>>>

In the context of inbreeding as the postulated driver of reproductive strategy evolution, the absolute and relative frequencies of alleles underlying pre-copulatory and post-copulatory inbreeding avoidance and polyandry will be affected not only by the magnitudes of positive indirect selection stemming from reduced inbreeding depression in females' offspring, but also by the magnitudes of direct negative selection on resulting phenotypes (i.e., the direct fitness costs of expressing each reproductive strategy). Empirical studies have demonstrated diverse costs of mating and mate choice, for example, including energetic costs of developing, maintaining, or enacting necessary physiologies [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014b]; increased risks of predation or disease stemming from increased mate-searching or mating [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]; increased risk of complete mating or fertilisation failure given extreme choosiness [@Kokko2013]; and risks of harm stemming from sexual conflict over fertilisation [e.g., @Rowe1994]. If the relative costs of pre-copulatory and post-copulatory inbreeding avoidance differ, then alleles underlying the less costly strategy might become fixed over generations, while alleles underlying the more costly strategy might go extinct, especially if their effects become redundant following evolution of the less costly strategy. Evolutionary dynamics will therefore depend on the inbreeding avoidance benefit of each strategy, which might also differ and vary in relation to the evolution of the other strategy. Dynamic models that track the frequencies of alleles underlying multiple potentially costly inbreeding avoidance strategies that might influence each other's evolution across generations, and that are enacted among relatives that themselves result from reproductive strategies and inbreeding depression expressed in previous generations, are consequently useful to understand and predict evolutionary outcomes.

@@

Phenotypic values of the three reproductive strategy traits each incur set costs that combine independently to increase the probability that a focal female will die before mating. Numerous forms and mechanisms of direct costs of reproductive strategies could potentially be hypothesised and modelled; the most appropriate formulation depends on the question posed (see Discussion). By allowing costs of the three strategies to be directly and independently controlled, our model facilitates direct comparison of evolution of pre-copulatory versus post-copulatory inbreeding avoidance given known cost asymmetries. Such costs on female survival probability are biologically reasonable. For example, polyandrous females that undertake increased mate searching or mating can experience increased predation risk [e.g., @Rowe1988; @Ronkainen1994; @Koga1998]. Females that express pre-copulatory choice can increase the risk of mortality due to harm caused by sexual conflict over mating, and also risk complete mating failure [which equates to pre-reproductive mortality in semelparous organisms; @Rowe1994; @Kokko2013]. Finally, females that express post-copulatory choice can pay up-front energetic costs, which might result in trade-offs with survival due to developing physiological or biochemical mechanisms needed to store sperm and successfully bias fertilisation [e.g., @Gasparini2011; @Tuni2013; @Fitzpatrick2014].

@@

To address how asymmetric costs associated with pre-copulatory and post-copulatory inbreeding avoidance and polyandry affect the long-term persistence of reproductive strategies, we quantify the change in $M_{a}$ and $F_{a}$ over generations in simulations where pre-copulatory inbreeding strategy was cost free ($c_{M}=0$) but post-copulatory inbreeding strategy was moderately costly ($c_{F}=0.02$), and vice versa. We compare evolutionary trajectories with those of a costly strategy in the absence of evolution of an alternative strategy (e.g., evolution of pre-copulatory inbreeding strategy when post-copulatory inbreeding strategy phenotype is fixed at zero, $F_{p}=0$). Previous modelling using similar genetic architecture suggests that a cost of $0.02$ imposes strong but not overwhelming direct negative selection on polyandry [@Duthie]. This value is therefore appropriate to illustrate the different evolutionary consequences that could result from asymmetrical costs (see Supporting Information for additional cost value combinations).
